---
title: "FianlProject"
author: "Rui Li"
date: "12/8/2020"
output: html_document
---
##read
```{r}
pacman::p_load(mlr, survival, survminer, tidyverse, readxl, lmtest, table1, nnet)
library(arsenal)
EDS=read_excel('COVID-19summary.xlsx')
```

##Modification: LOS is not in the same format
```{r}
## change LOS (?d?h) to total hour.
#label data with "day"
EDS = EDS %>%
  mutate(label = ifelse(str_detect(LOS, "d"), 1, 0))
#delete 'd' and 'h'
EDS.1 = EDS %>%
  mutate(LOS = str_remove(LOS, "h")) %>%
  filter(label==1) %>%
  separate(LOS, into = c("day","hour"), sep= "d", remove = F) %>%
  mutate(hour = as.numeric(hour),
          day = as.numeric(day)) %>%
  mutate(hour = ifelse(is.na(hour),0, hour)) %>%
  mutate(hour_total=day*24 + hour) %>%
  select(-day, -hour)
  

EDS.0 = EDS %>%
  filter(label==0) %>%
  mutate(hour_total = as.numeric(str_remove(LOS, "h")))

#merge
COVID = rbind(EDS.1,EDS.0)
              
```

##Data management
```{r}
COVID = COVID %>%
##Recode Age variable as a factor variable and label it
  mutate(Age = as.numeric(Age),
         Age_f=case_when(Age < 60 ~ "0",
                         Age >= 60  ~ "1"),
##Recode LOS variable as a factor variable
         ER_retention=case_when(hour_total < 2 ~ "0",
                                hour_total >=2 ~ "1"))
##delete the missing value (optional)
    ##filter(is.na()==FALSE) %>%
COVID = COVID %>% 
  na.omit()
```

```{r}
print(head(COVID, 10))
```

##table 1
```{r}
COVID.T = COVID %>%
##Recode Age variable as a factor variable and label it
  mutate(Age = as.numeric(Age),
         Age_f = case_when(Age < 60 ~ "No Elderly",
                         Age >= 60  ~ "Elderly"),
##Recode LOS variable as a factor variable
         ER_retention = case_when(hour_total < 2 ~ "No Retention",
                                hour_total >=2 ~ "Retention"),
          Gender = case_when(Sex == 1 ~ "Male",
                          Sex == 0 ~ "Female"),
          EPI = case_when(EPI == 1 ~ "Post-epidemic",
                          EPI == 0 ~ "Pre-epidemic"))

label(COVID.T$Age_f) = "Elderly"
label(COVID.T$System) = "System of human body"
label(COVID.T$ER_retention) = "Emergency room retention"
table1(~ Age_f + Gender + System + Outcome + ER_retention|EPI, COVID.T)
```

## Multivariate linear regression: LOS ~ Age + Sex + System
```{r}
COVID = COVID %>%
  mutate(System = as.factor(System) )

mod.mlr = lm( hour_total ~ EPI + Age + Sex + System, data= COVID)
summary (mod.mlr)

confint(mod.mlr, level = 0.95)
## Interpretation: After adjusting for covariates, the mean LOS of patients during COVID-19 increased 13.5 h compared to pre-epidemic.
```

##Logistic regrssion -- retention
```{r}
#logistic model
COVID$ER_retention = as.numeric(COVID$ER_retention)
Logit <- glm(ER_retention ~ EPI, data=COVID, family="binomial")
summary(Logit)

#calculate and print ORs and 95% CIs  
OR<-exp(cbind(OR = coef(Logit), confint(Logit)))
OR
#Interpretation: the odds of retention is 5.7 times higher (95% CI 4.4-7.4)in patients admitted during the COVID-19 post-epidemic compared to pre-epidemic.
```
```{r}
#logistic model -adjusted
COVID$ER_retention = as.numeric(COVID$ER_retention)
Logit.a <- glm(ER_retention ~ EPI+ Age + Sex + System, data=COVID, family="binomial")
summary(Logit.a)

#calculate and print ORs and 95% CIs  
OR<-exp(cbind(OR = coef(Logit.a), confint(Logit.a)))
OR
#Interpretation: the odds of retention is 5.7 times higher (95% CI 4.4-7.4)in patients admitted during the COVID-19 post-epidemic compared to pre-epidemic.
```

## Cox Proportional-Hazards Model (death, discharge, transfer = event)
## EVENT--Leaving the ER
```{r}
## Figure
hist(COVID$hour_total)

## Treat 'transfer', 'discharge', 'death' as event, 'Obeservation' as censoring.
      ## recode variable- event
 COVID1 = COVID %>%
   mutate(event = case_when(Outcome %in% c('transfer', 'discharge', 'death') ~ "1",
                           Outcome %in% c('Observation') ~ "0"),
         event = as.numeric(event),
         EPI = factor(EPI),
         Sex = factor(Sex),
         System = factor(System),
         EPI = case_when(EPI == 1 ~ "Post-epidemic",
                          EPI == 0 ~ "Pre-epidemic"))
      ## examine the recoding  
 table(COVID1$Outcome, COVID1$event)
 
##Modling
 mod.cox1.unadj = coxph(Surv(hour_total, event) ~ EPI, COVID1)
 summary(mod.cox1.unadj)
##Interpretation: Those who were admitted to the ER during post-epidemic had a 3.9 times higher hazard of retention than pre-epidemic.

 ## Adjusting 
 mod.cox1 = coxph(Surv(hour_total, event) ~ EPI + Age + Sex + System, COVID1)
 summary(mod.cox1)
##Interpretation: Those who admitted to ER during the epidemic had a 4.3  times higher hazard of retention than pre-epidemic.
 
##Compare models
lrtest(mod.cox1.unadj, mod.cox1) 

###The likelihood ratio test indicates that the model adjusting confounders significantly improves fit compared to the model with just epidemic (p=3.6 * e-08). 
 ##### ERROR: problem of plot
##class(COVID$Sex)
##mod2 = coxph(Surv(hour_total, event) ~ EPI + Age + Sex + System, COVID)
##summary(mod2)
##ggadjustedcurves(mod.cox2, data = COVID1, method = "marginal", variable = "EPI")

##KM
library(ggfortify)
library(survival)
fit <- survfit(Surv(hour_total, event) ~ EPI, data = COVID1)

## ggsurvplot (xlab = "Hours")
ggsurvplot(fit, data = COVID1, risk.table = TRUE, conf.int=TRUE, tables.theme = clean_theme(), xlab = "Hours") 
fit


COVID1$day_total = COVID1$hour_total/24 ## Add day variable
fit2 <- survfit(Surv(day_total, event) ~ EPI, data = COVID1)

## ggsurvplot (xlab = "Days")
ggsurvplot(fit2, data = COVID1, risk.table = TRUE, conf.int=TRUE, tables.theme = clean_theme(), xlab = "Days") 
fit2

## Day < 3
COVID.D.F = filter(COVID1, day_total < 3)
fit3 <- survfit(Surv(day_total, event) ~ EPI, data = COVID.D.F)
## ggsurvplot (xlab = "Days")
ggsurvplot(fit3, data = COVID.D.F, risk.table = TRUE, conf.int=TRUE, tables.theme = clean_theme(), xlab = "Days") 
fit3
survdiff(Surv(day_total, event) ~ EPI,data=COVID.D.F)
```

## Cox Proportional-Hazards Model (discharge, transfer = event)
## EVENT--Leaving the ER
```{r}
 ## Treat 'transfer', 'discharge' as event.
  COVID2 = COVID %>%
    filter(Outcome!= 'death') %>%
   mutate(event = case_when(Outcome %in% c('transfer', 'discharge') ~ "1",
                           Outcome %in% c('Observation') ~ "0"),
         event = as.numeric(event),
         EPI = factor(EPI),
         Sex = factor(Sex),
         System = factor(System))
       ## examine the recoding  
 table(COVID2$Outcome, COVID2$event)
       ## Figure

##Modling
 mod.cox2.unadj = coxph(Surv(hour_total, event) ~ EPI, COVID2)
 summary(mod.cox2.unadj)

 ##Interpretation: Those who admitted to ER during the epidemic had a 4.3  times higher hazard of retention than pre-epidemic.

 1/0.1977
 1/0.2682
 
## Adjusting 
 mod.cox2 = coxph(Surv(hour_total, event) ~ EPI + Age + Sex + System, COVID2)
 summary(mod.cox2)
##Interpretation: Those who admitted to ER during the epidemic had a 4.3  times higher hazard of retention than pre-epidemic.
 1/0.1810    
 1/0.2480
##Compare models
lrtest(mod.cox2.unadj, mod.cox2) 

###The likelihood ratio test indicates that the model adjusting confounders significantly improves fit compared to the model with just epidemic (p=1.9 * e-11). 
```

## Cox Proportional-Hazards Model (Death = event)
## EVENT--Death
```{r}
 ## Treat 'death' as event.
  COVID3 = COVID %>%
   mutate(event = case_when(Outcome %in% c('death') ~ "1",
                           Outcome %in% c('transfer', 'discharge', 'Observation') ~ "0"),
         event = as.numeric(event),
         EPI = factor(EPI),
         Sex = factor(Sex),
         System = factor(System))
       ## examine the recoding  
 table(COVID3$Outcome, COVID3$event)
       
 
##Modling
 mod.cox3.unadj = coxph(Surv(hour_total, event) ~ EPI, COVID3)
 summary(mod.cox3.unadj)

 ##Interpretation: No significant difference was found. Those who admitted to ER during the epidemic had a 1.4  times higher hazard of death than pre-epidemic.

## Adjusting 
 mod.cox3 = coxph(Surv(hour_total, event) ~ EPI + Age + Sex + System, COVID3)
 summary(mod.cox3)
##Interpretation: No significant difference was found. Those who admitted to ER during the epidemic had a 1.5  times higher hazard of death than pre-epidemic.
 
##Compare models
lrtest(mod.cox3.unadj, mod.cox3) 

###The likelihood ratio test indicates that the model adjusting confounders significantly improves fit compared to the model with just epidemic (p=1.4 * e-8). 
```

## Multinominal logistic regression 
```{r}
##Circulatory & Cardiovascular system as the reference
mod.mlog = multinom ( System ~ EPI + Age + Sex , data = COVID)
summary(mod.mlog)
exp(coef(mod.mlog))
exp(confint(mod.mlog, level=0.95))

##set Respiratory system disease as the reference
COVID.SYS = COVID %>%
  mutate(System = factor(System),
         System = relevel(System, ref = 'Respiratory system'))
mod.mlog2 = multinom ( System ~ EPI + Age + Sex , data = COVID.SYS)
summary(mod.mlog2)
exp(coef(mod.mlog2))
exp(confint(mod.mlog2, level=0.95))
##Interpretation: The odds of being diagnosed with Circulatory & Cardiovascular system diseases relative to Respiratory system disease is 0.67 times lower for post-epidemic ER admission patients versus pre-epidemic ER admission patients, holding all other variables constant. 
##Interpretation: The odds of being diagnosed with Muscular & Skeletal System diseases relative to Respiratory system disease is 3.0 times higher for post-epidemic ER admission patients versus pre-epidemic ER admission patients, holding all other variables constant. 
##Interpretation: The odds of being diagnosed with Nervous system diseases relative to Respiratory system disease is 1.5 times higher for post-epidemic ER admission patients versus pre-epidemic ER admission patients, holding all other variables constant. 
##Interpretation: The odds of being diagnosed with Renal & Urinary system diseases relative to Respiratory system disease is 3.7 times higher for post-epidemic ER admission patients versus pre-epidemic ER admission patients, holding all other variables constant. 
##Interpretation: The odds of being diagnosed with Reproductive system diseases relative to Respiratory system disease is 1.9 times higher for post-epidemic ER admission patients versus pre-epidemic ER admission patients, holding all other variables constant. 
```

